### 2020-06-13 appendix/tikz2pdf.tex

\documentclass{article}
\usepackage[pdftex,active,tightpage]{preview}
\usepackage[utf8]{inputenc}
\usepackage[skins]{tcolorbox}
\usepackage{
tikz,
enumitem,
xcolor,
pgfplots
}
\begin{document}

  \tikzstyle{phase} = [rectangle, rounded corners, minimum width = 3cm, minimum height = 2cm, text centered, text width = 4cm, draw = black, thick, fill=white]
  \tikzstyle{exc} = [rectangle, rounded corners, minimum width = 3cm, minimum height = 2cm, text centered, text width = 4cm, draw = black, thick, fill=gray!10]
  \tikzstyle{stage} = [rectangle, minimum height = 1cm, minimum width = 2cm, text centered, draw = black, very thick]
  \tikzstyle{stage2} = [rectangle, minimum height = 1cm, minimum width = 3.5cm, text centered, draw = black, very thick]
  \tikzstyle{arrow} = [ultra thick,->,>=stealth]

\begin{preview}
%% TIKZ_CODE %%
\end{preview}

\end{document}